% 21:09 04/04/2023
\begin{appendices}
\renewcommand{\chaptername}{Appendix A}
\renewcommand{\thechapter}{A}
\chapter{}
The GitHub repository for the project can be found at the URL: \\
\href{https://github.com/G00378925/caladium}{https://github.com/G00378925/caladium}

\renewcommand{\chaptername}{Appendix B}
\renewcommand{\thechapter}{B}
\chapter{}
\label{sec:kanban_table}

\begin{tabular}{|c|p{14cm}|}
    Card No. & Task \\
    \hline
    1 & The server needs to be able to allocate jobs to the sandbox instances. \\
    \hline
    2 & Dashboard needs to be able to add sandbox instances through the UI. \\
    \hline
    3 & The server needs to be able to upload files to be run. \\
    \hline
    4 & Get Docker setup and write a Dockerfile. \\
    \hline
    5 & The single-page application needs to be able to generate HTML dynamically. \\
    \hline
    6 & Add a preferences menu to the dashboard so admins can update settings. \\
    \hline
    7 & Clients need to be able to pass files to the server to be scanned. \\
    \hline
    8 & The dashboard needs to be able to ping workers to see if they are alive. \\
    \hline
    9 & Add code to check if system calls are malicious. \\
    \hline
    10 & Rewrite syscall analysis in Python because it is too slow in Racket. \\
    \hline
    11 & Add piecharts and barcharts to the index page on the dashboard to display statistics. \\
    \hline
    12 & Make variable names consistent. \\
    \hline
    13 & When a user installs the Caladium client,
    the client must request the provisioning token. \\
    \hline
    14 & GitHub Action to automatically run the test suite
    when a commit is pushed to the repository. \\
    \hline
    15 & Add an uninstall function to the Caladium client. \\
    \hline
    16 & Add the static analysis code to the worker. \\
    \hline
    17 & Setup GitHub pages for the promotional page. \\
    \hline
    18 & Allow administrators to be able to set custom IPs
    for the server when building clients. \\
    \hline
    19 & Add a label to the main page of the client,
    showing the currently scanned directory. \\
    \hline
    20 & Fix the auto-provisioning bug where it doesn't
    provision after inputting the profile JSON. \\
\end{tabular}

\renewcommand{\chaptername}{Appendix C}
\renewcommand{\thechapter}{C}
\chapter{}

The following are instructions to build each component of Caladium:
First, retrieve the latest build of Caladium, you can do that with this command:
\texttt{git clone https://github.com/G00378925/caladium.git}

Python is a requirement for all projects, you can
\href{https://www.python.org/downloads/}{download and install, the latest version for your system here}

% <!-- --><!-- --><!--        --><!-- --><!-- -->
% <!-- --><!-- --><!--        --><!-- --><!-- -->
% <!-- --><!-- --><!-- Client --><!-- --><!-- -->
% <!-- --><!-- --><!--        --><!-- --><!-- -->
% <!-- --><!-- --><!--        --><!-- --><!-- -->
\section{Client}
You must make sure you are on Microsoft Windows
for this part as \textbf{iexpress.exe} is required.
Run the following commands in \textbf{Command Prompt},
making sure you are in the Caladium root directory.

\begin{lstlisting}
cd client
python -m pip install -r requirements.txt

rem You will be prompted to enter the IP address of the Caladium server

build.cmd
\end{lstlisting}

This will result in a \texttt{caladium-setup.exe} in the \texttt{dist} directory,
this is the installer that will be given to users.
% <!-- --><!-- --><!--        --><!-- --><!-- -->
% <!-- --><!-- --><!--        --><!-- --><!-- -->
% <!-- --><!-- --><!-- Server --><!-- --><!-- -->
% <!-- --><!-- --><!--        --><!-- --><!-- -->
% <!-- --><!-- --><!--        --><!-- --><!-- -->
\section{Server}
To build and run the server, enter the following commands into your terminal,
these should work on Linux, macOS and Windows.

You need the address of your CouchDB instance this needs
to be put in the environmental variable \texttt{COUCHDB\_CONNECTION\_STR}.
If on Windows use
\texttt{set COUCHDB\_CONNECTION\_STR=admin:root@0.0.0.0:5984},
if on Linux or macOS use
\texttt{export COUCHDB\_CONNECTION\_STR=admin:root@0.0.0.0:5984}.
With \texttt{admin:root@0.0.0.0:5984} being the CouchDB connection string.

Make sure to swap out `python3` with `python` if executing on Windows.
\begin{lstlisting}
cd server/src

python3 -m pip install flask requests
python3 __main__.py
\end{lstlisting}
The address of the instance will be printed on the terminal.

\subsubsection{Using Docker}
Instead of having to set up an environment manually,
you can use the supplied Dockerfile
to set up a reproducible environment,
replace the \texttt{0.0.0.0:5984} in the commands below like above.

Make sure you have Docker installed on your system,
and type the following into your terminal.

\begin{lstlisting}
cd server
sudo docker build --tag caladium .
sudo docker run -e COUCHDB_CONNECTION_STR=^
    "http://admin:root@0.0.0.0:5984" -p 80:8080 caladium
\end{lstlisting}
% <!-- --><!-- --><!--                  --><!-- --><!-- -->
% <!-- --><!-- --><!--                  --><!-- --><!-- -->
% <!-- --><!-- --><!-- Sandbox Analysis --><!-- --><!-- -->
% <!-- --><!-- --><!--                  --><!-- --><!-- -->
% <!-- --><!-- --><!--                  --><!-- --><!-- -->
\section{Sandbox Analysis}
To set up the analysis service on a computer, you must first
\href{https://download.racket-lang.org/}{install Racket} and
\href{https://sandboxie-plus.com/}{Sandboxie}.
Create a \texttt{SysinternalsSuite} directory at the root of your drive, this is usually \texttt{C:\\SysinternalsSuite}.
Place \texttt{Procmon64.exe} in there, you can
\href{https://learn.microsoft.com/en-us/sysinternals/downloads/procmon}{download it here}.

Alternatively, you can use these commands to download Procmon to that location.
\begin{lstlisting}
mkdir /l %SystemDrive%\SysinternalsSuite 2>null
curl https://live.sysinternals.com/Procmon64.exe ^
    --output %SystemDrive%\SysinternalsSuite\Procmon64.exe
\end{lstlisting}

ClamAV is required for the static analysis,
I am using a distribution of ClamAV called ClamWin,
\href{https://clamwin.com/content/view/18/46/}{which can be downloaded here},
ClamWin will automatically download the latest malware definitions.

Run a new Command Prompt as administrator, and run \texttt{cd sandbox \\
\&\& start\_sandbox.cmd}.
The analysis service will now poll for tasks, in the case of it crashing it will automatically restart.
Note the IP address and port of the service,
in the format \texttt{0.0.0.0:8080}
and add it to the administrator dashboard.

\end{appendices}
